---
title: "Counter"
author: "Andreas M. Brandmaier"
date: "12/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup

First, we load the package and we set a common theme for our plots later.

```{r load, echo=TRUE, eval=TRUE, message=FALSE}
library(bnnlib)
mytheme <- ggthemes::theme_hc()
```

## Generate Data

Generate some sequences with a counting task. 

```{r}
set.seed(234) # for all R commands
setRandomSeed(234) # for all C++ commands

seqset <- SequenceSet()

max.num.spikes <- 10
num.sequences <- 100

for (i in 1:num.sequences) {
  ln <- sample(20:150,1)
  input <- rep(0,ln)
  num.spikes <- sample(1:max.num.spikes,1)
  input[sample(1:ln,num.spikes,replace=FALSE)]<-1
  target <- cumsum(input)#/max.num.spikes
  seq <- toSequence( data.frame(input,target),1,2)
  SequenceSet_add_copy_of_sequence(seqset, seq)
}
```


## Training

Create a recurrent neural network and train for several epochs.

```{r training, echo=TRUE, message=FALSE}

network = LSTMNetwork(1,1,1, bnnlib::TANH_NODE, bnnlib::LINEAR_NODE)
#network = RecurrentNetwork(1,10,1)

trainer = ImprovedRPropTrainer(network)
#trainer = RPropTrainer(network)  
Trainer_train__SWIG_0(trainer, seqset, 400)

```

Plot the prediction error over training epochs.

```{r plottraining}
values <- getTrainingerror(trainer)

library(ggplot2)
ggplot(data.frame(x=1:length(values),values),aes(x=x,y=values))+
  geom_line()+
	mytheme+
  xlab("Epochs")+ ylab("Error")+
  ggtitle("Trainingset Error")+
  scale_y_log10()
```

Get activations of the first sequence in the sequence set and plot

```{r}
seq <- SequenceSet_get(seqset, 0)
x = getActivations(network, seq)


nms <- rep(NA, ncol(x))
for (i in 1:ncol(x))
  nms[i] = Network_get_node_name(network, i-1)

nms <- make.unique(nms)

xlong <- reshape2::melt(x)
xlong$Var2 <- factor(xlong$Var2, labels = nms)
ggplot(data=xlong, aes(x=Var1,y=value,group=Var2,col=Var2))+
  geom_line() + mytheme
```

Let's plot the same thing for only the constant error carousels (CEC):

```{r}

xlongfilt <- xlong[substr(xlong$Var2,1,3)=="CEC",]
ggplot(data=xlongfilt, aes(x=Var1,y=value,group=Var2,col=Var2))+geom_line()+
  mytheme

```

Check match between targets (true values) and outputs (network predictions). To this end, we write a little helper function to create plots.

```{r}
plot_to<-function(index) {
  seq <- SequenceSet_get(seqset, index-1)
  outputs <- getOutputs(network, seq)
  targets <- getTargets(seq)
  
  df <- data.frame(vals=c(outputs,targets),time=rep(1:length(outputs),2),
                   grp=factor(rep(c("output","target"),each=length(outputs))))
  
  pl <- ggplot(data=df, aes(x=time,y=vals,group=grp,col=grp))+
    geom_line()+mytheme
  
  (pl)
}
```

```{r}
plot_to(1)
plot_to(2)
plot_to(5)
plot_to(20)
```

## Network Topology

Finally, plot the network topology with learnt weights:

```{r results="asis"}
dot.lang <- Network_to_dot_graph(network)
library("DOT")
DOT::dot(dot.lang)
```