---
title: "dynamic.Rmd"
author: "Andreas M. Brandmaier"
date: "3/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

dyn.load(paste("../bnnlib", .Platform$dynlib.ext, sep=""))
source("../bnnlib.R")
cacheMetaData(1)
```

## Feedforward Neural Network

Create a feed-forward neural network with 3 inputs, 25 hidden units, and 1 output. The output node is linear.

```{r}
TANH_NODE = 1
LINEAR_NODE = 3

net <- NetworkFactory_createFeedForwardNetwork(2,4,3, LINEAR_NODE)
```

## Randomly remove some connections

Here, we first count the total number of trainable weights and then remove weights with a 15% chance.
```{r}
Network_get_num_trainable_weights(net)

no_removed = Network_randomly_remove_connections(net, 0.15)
cat("Connections removed: ", no_removed,"\n")
```

Again, let's count the total number of weights and remove some more (with 10%).

```{r}

Network_get_num_trainable_weights(net)

no_removed = Network_randomly_remove_connections(net, 0.10)

Network_get_num_trainable_weights(net)

```

Introduce a new node

```{r}
# get references to the three layers (0=bias); 
# this enumeration is true only for feedforward networks from NetworkFactors
bias_layer <- Network_get_ensemble(net, 0);
input_layer <- Network_get_ensemble(net, 1);
hidden_layer <- Network_get_ensemble(net, 2);
output_layer <- Network_get_ensemble(net, 3);
# create new node and add it to the hidden layer
new_node <- NodeFactory_createNode(TANH_NODE)
Node_name_set(new_node,"New Node")
Ensemble_add_node(hidden_layer, new_node)
```

Plot the network architecture:


```{r results="asis"}
dot.lang <- Network_to_dot_graph(net)
require("DOT")
DOT::dot(dot.lang)
```

```{r}
# add connections from new node to output layer
for (i in 1:Ensemble_get_num_nodes(output_layer)) {
  output_node = Ensemble_get_node(output_layer, i-1);
 con= Network_connect__SWIG_0(new_node, output_node, TRUE);
}

Network_connect__SWIG_0(Ensemble_get_node(bias_layer,0), new_node, TRUE)
```


Now, let's count again the number of trainable weights:

```{r}
Network_get_num_trainable_weights(net)
```

Now, we add connections from all input layer nodes to the new node.

```{r}
# add connections from input layer to new node
for (i in 1:Ensemble_get_num_nodes(input_layer)) {
  input_node = Ensemble_get_node(input_layer, i-1);
  Network_connect__SWIG_0(input_node, new_node, TRUE);
}
# TODO: also add weight from bias node?
```

Again, let's count again the number of trainable weights:

```{r}
Network_get_num_trainable_weights(net)
```


```{r results="asis"}
dot.lang <- Network_to_dot_graph(net)
require("DOT")
DOT::dot(dot.lang)
```