---
title: "FeedforwardNetwork"
author: "Andreas M. Brandmaier"
date: "12/11/2019"
output:
  html_document:
    df_print: paged
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Feedforward Neural Network

Load the library first
```{r message=FALSE}
library(bnnlib)
```
Create a feed-forward neural network with 3 inputs, 25 hidden units, and 1 output. The output node is linear.

```{r}
net <- NetworkFactory_createFeedForwardNetwork(3,25,1, 
                              bnnlib::LINEAR_NODE)
```

Load the marketing dataset from the `datarium` package.

```{r}
data("marketing", package = "datarium")
head(marketing, 4)
```

Convert data to Sequence format:

```{r}
sequence_set <- SequenceSet()
seq <- toSequence(marketing, 1:3, 4)
SequenceSet_add_copy_of_sequence(sequence_set,seq)

cat("Sequences created: ",SequenceSet_size(sequence_set),"\n")
cat("Input size",SequenceSet_input_size_get(sequence_set),"\n")
cat("Target size",SequenceSet_target_size_get(sequence_set),"\n")

```

Initialize training algorithm

```{r message=FALSE}
bp <- ImprovedRPropTrainer(net)

cat("Using Trainer: ",Trainer_get_name(bp),"\n")
```

Start training

```{r}
iterations <- 200
steps.per.iteration <- 1
err <- rep(NA, iterations+1)
err[1] <- Network_evaluate_training_error__SWIG_0(net, sequence_set)
for (i in 1:iterations) {
  Trainer_train__SWIG_0(bp, sequence_set, steps.per.iteration)
  err[i+1] <- Network_evaluate_training_error__SWIG_0(net, sequence_set)
}
```

With ggplot2, we can plot the training set error over iterations:

```{r warning=FALSE, message=FALSE}
require(ggplot2)
ggplot(data=data.frame(step=1:(iterations+1),err),aes(x=step,y=err))+
  geom_point()+ geom_line()+
  geom_smooth()+
  theme_minimal()
```

Inspect output:

```{r}

outputs <- getOutputs(net, seq)
targets <- getTargets(seq)
ggplot(data=data.frame(outputs=outputs, targets=targets), aes(y=outputs,x=targets))+
  geom_point()+
  theme_minimal()+
  xlab("Targets")+ylab("Predictions")+
  geom_smooth()+
  ggtitle("Sales")

```
