---
title: "Mean Spike Distance"
author: "Andreas M. Brandmaier"
date: "1/14/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is a demonstration of the mean spike distance task by Gers et al. (2002). The task is about learning a representation of time, which involves counting discrete time steps, reading out the counter and also learning to reset the timer when an external stimulus is present.

First, we load the library and some plotting libraries and set a random seed for all bnnlib internals.

```{r}
library(bnnlib)
library(ggplot2)
library(gridExtra)
library(tictoc)
setRandomSeed(92357)
```

We create the network from Gers02a with LSTM hidden layer (including peepholes) with 1 output and 1 input.
```{r}
#network = NetworkFactory_create_gers02a()
network = LSTMNetwork(1,1,1, TANH_NODE, LINEAR_NODE)
```

Generate data
```{r}
set.seed(4598234)
n <- 100
interval.lengths <- 1:5
#interval.lengths <- 1:10
sequence_set <- SequenceSet()

for (k in 1:10) {
intervals <- sample(interval.lengths, size=n, replace=TRUE)
input_seq <- c(1,rep(0, sum(intervals)))
target_seq <- rep(0, sum(intervals)+1)
input_seq[cumsum(intervals)+1] <- 1
target_seq[cumsum(intervals)+1] <- intervals
sequence <- toSequence(data.frame(input_seq, target_seq),1,2)
SequenceSet_add_copy_of_sequence(sequence_set, sequence)
}
```

## Training

```{r}
#trainer <- ADAMTrainer(network)
trainer <- ImprovedRPropTrainer(network)
#trainer <- ARPropTrainer(network)


tic()
Trainer_add_abort_criterion__SWIG_0(trainer, ConvergenceCriterion(1e-3),10)
Trainer_train2(trainer, sequence_set, 500)
toc()


plotTrainingerror(trainer) + theme_light()
```

## Predictions

Plot the network predictions zoomed in to the first 25 steps of the sequence.
```{r}
plotPredictions(network, sequence) + theme_light()
```

## Investigate activations

Plot all activations for the entire sequence.

```{r}
plotActivations(network, sequence) + theme_light() +   coord_cartesian(xlim=c(0,25))

plotActivations(network, sequence,"CEC") + theme_light() +   coord_cartesian(xlim=c(0,25))

plotActivations(network, sequence) + theme_light() +   coord_cartesian(xlim=c(0,25))
```