---
title: "Mushroom"
author: "Andreas M. Brandmaier"
date: "12/12/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Classification with Feedforward Networks

Example from the Proben1 dataset.

```{r pressure, echo=FALSE, message=FALSE}
library(bnnlib)
```

Download the mushroom dataset in Proben1 file format directly from github:

```{r}
temp <- tempfile()
download.file("https://raw.githubusercontent.com/jeffheaton/proben1/master/mushroom/mushroom1.dt",temp)
dataset <- SequenceSet_load_from_proben_file(temp)
```

Create a network with 125 inputs and 2 outputs. Create a training algorithms for this network.

```{r}
network <-NetworkFactory_createFeedForwardNetwork(125, 20, 2, bnnlib::SIGMOID_NODE)
Network_reinitialise(network)
#cg <- CGTrainer(network)
cg <- ImprovedRPropTrainer(network)
```

Train the network for some steps

```{r}
Trainer_add_abort_criterion__SWIG_0( cg, ConvergenceCriterion(), 5 )
Trainer_train__SWIG_1(cg, DataSet_get_training_set(dataset), DataSet_get_validation_set(dataset), 20)
```

Evaluate network performance using a winner-takes-all classification strategy. This is the total number of wrong classifications:

```{r}
error_function = WinnerTakesAllErrorFunction();
Network_evaluate_error__SWIG_1(network, DataSet_get_test_set(dataset), error_function)
```

Inspect network predictions

```{r}
test_sequence_set <- DataSet_get_test_set(dataset)
lngth <- SequenceSet_size(test_sequence_set)

result <- matrix(NA, nrow=lngth, ncol=2)
truth <- matrix(NA, lngth, ncol=2)
for (i in 1:lngth) {
    sq<-SequenceSet_get(test_sequence_set, i-1)
  Network_activate(network, sq)
  x<- Network_get_output(network,0)
  result[i,] <- .Call('R_swig_toValue', x, package="bnnlib")
  x<-Sequence_get_target(sq, 0)
  truth[i,] <- .Call('R_swig_toValue', x, package="bnnlib")
  
}
predictions <- apply(result, 1, which.max)
truths <-  apply(truth, 1, which.max)

confusion.matrix <- table(truths, predictions)



knitr::kable(confusion.matrix, row.names=TRUE)
```
