---
title: "Stochastic Feedforward Network"
author: "Andreas M. Brandmaier"
date: "SEP/03/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bnnlib)
```

## Stochastic Gain

In the computational model of ageing by Li et al., sigmoid cells with stochastic gain implement ageing networks. As networks grow older, the stochastic gain is decreased. The young network has an average gain of .8 and the old network an average gain of .3. 

```{r}

gain = .8 # young gain = .8; old gain = .3
bias = +1
stochastic_range = .2
input_signal = 4
num.trials = 1000

# specify the network topology
num_hid = 5 # number of hidden units
in_size = 1 # number of input units
out_size = 1 # number of output units
num_layers = 1 # number of hidden layers (do not count input and output layer)
LINEAR_NODE = 3 # activation type for output layer (3=linear)
```

Let's create the network and systematically vary the connection strength from the input to the hidden layer nodes.

```{r}
# create the network using the factory method from bnnlib
net <- NetworkFactory_createStochasticFeedForwardNetwork(
  in_size, num_layers, num_hid,out_size, LINEAR_NODE,
  gain, bias, stochastic_range)

# manipulate connection
nodenames <- getNodeNames(net)
input_node <- Network_get_node_with_name(net, nodenames[2])
nc = Node_get_num_outgoing_connections(input_node)
weights <- rep(NA, nc)
for (i in 1:nc) {
  conn = Node_get_outgoing_connection(input_node, i-1)
  w <- (i/nc) 
  weights[i] <- w
  cat("Set weight to ",w,"\n")
  Trainable_weight_set(conn, w );
}
```

Next, we create a training data with fixed input that will be used to assess the stochasticity given a fixed input over multiple trials.

```{r}
# create a dataset with constant input
x <- data.frame(matrix(input_signal,nrow=num.trials,ncol=2))
seq=toSequence(x,1,2)

get.activation.long <- function(net, seq) {
  # activate the network with the inputs and
  # filter only the hidden unit activations
  act<-data.frame(getActivations(net, seq))
  colnames(act) <- getNodeNames(net)
  filter <- startsWith(getNodeNames(net),"Stochastic")
  act <- act[,filter]
  # convert activation data from wide format to long format
  act <- cbind(act, ID=1:nrow(act))
  longdat <- tidyr::gather(act,"key","value", 1:(ncol(act)-1) )
  # return result
  return(longdat)
}

# get activation
data.young <- get.activation.long(net, seq)
```

Now, we manipulate the gain parameter for all hidden layer units from 0.8 (young) to 0.3 (old).

```{r}
# now reset the gain parameter for one node
nodename <- unique(data.young$key)
for (name in nodename) {
  cat("Setting ",name, ":")
  node = Network_get_node_with_name(net,name)
  x = StochasticSigmoidNode_gain_get(node)
  cat (" gain was at ",x,"\n")
  StochasticSigmoidNode_gain_set(node, 0.3)
}

```

Let's collect all activations again.

```{r}

# get activation
data.old <- get.activation.long(net, seq)

# merge activations
data.all <- cbind( rbind(data.young, data.old), Agegroup=rep(c("Young","Old"),each=nrow(data.old) ) )
data.all$key <- factor(data.all$key,levels = rev(unique(data.all$key)),labels=paste0("W=",weights))

```

Now, let's plot the histograms of network activation with color representing the age groups.

```{r}
library(ggplot2)
gp <- ggplot(data=data.all, aes(x=value,group=Agegroup,fill=Agegroup))+
  #geom_density()+
  geom_histogram()+
  facet_wrap(~key)+
  scale_fill_manual(values=c("#c8c8dc","#f0d246"))+
  ggthemes::theme_clean()+
  xlab("Activation")+ ylab("Frequency")

plot(gp)
```

Finally, let's have a look at the network topology.

```{r results="asis"}
 dot.lang <- Network_to_dot_graph(net)
 require("DOT")
 DOT::dot(dot.lang)
```
